---
title: "Lab 3"
author: 
  - Makayla Whitney
  - Joe Swinehart
  - Janette Avelar
  - David Fainstein
date: "2/8/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide', messages = FALSE)

#loading libraries needed for Lab 3
library(ggplot2)
library(here)
library(tidyverse)
library(lubridate)
library(waffle)
library(dplyr)
library(directlabels)
library(gghighlight)
library(scales)
library(hexbin)

#Importing files
imported_honey_production <- read.csv(here("data/honeyproduction.csv"))
census_regions <-  read_csv(here("data/us census bureau regions and divisions.csv"))

```

```{r making data friendly}
#Pairing down table to a more manageable amount of columns; one mutation
purifiedhoney <- imported_honey_production %>% 
  select(state, year, totalprod) %>% 
  mutate(prodsimpler = totalprod/1000)

#One column name reassignment
census_regions <- census_regions %>% 
  rename(state = `State Code`)
```

\n Vizualization #1: Honey Production Over Time \n\n

```{r visualize honey production over time}

ggplot(purifiedhoney, aes(year, prodsimpler)) +
  #inserting all lines first
  geom_line(aes(group=state),
            color = "gray80") +
  #then 3 geom_lines to highlight OR, WA, and CA
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "#60D838") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "#00A1FE") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "#FF634D") +
  #Adjusting Y-axis to be more conducive to viewing
  scale_y_log10()+
  #Adding in appropriate indications and annotations
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "The average honey bee's wings flap over 183 times per second.")+    
  annotate(geom = "point", x = 1998, y = 37350, colour = "#FF634D", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "#00A1FE", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "#60D838", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")

```
\n 

Vizualization #2: Alteration for Accessibility #1 
\n\n

```{r colorblind friendly 1 Duetanomaly}

#Replication of code block above with one change: colors changed for accessibility to people with red-green colorblindness
ggplot(purifiedhoney, aes(year, prodsimpler)) +
  geom_line(aes(group=state),
            color = "gray80") +
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "#E5B33F") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "#8EA7E7") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "#FFE26A") +
  scale_y_log10()+
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "A queen bee can sting multiple times without dying.")+    
  annotate(geom = "point", x = 1998, y = 37350, colour = "#FFE26A", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "#8EA7E7", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "#E5B33F", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")
```
\n\n 

Vizualization #3: Alteration for Accessibility #2 \n\n

```{r colorblind friendly 2 Protanomaly}

#Replication of code block above with one change: colors changed for accessibility to people with a different type of red-green colorblindness
ggplot(purifiedhoney, aes(year, prodsimpler)) +
  geom_line(aes(group=state),
            color = "gray80") +
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "#FFD960") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "#A2B6EB") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "#D5A238") +
  scale_y_log10()+
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "A typical Bee colony consists of 30,000 to 60,000 bees.")+    
  annotate(geom = "point", x = 1998, y = 37350, colour = "#D5A238", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "#A2B6EB", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "#FFD960", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")
```

\n\n 

Vizualization #4: Alteration for Accessibility #3 \n\n

```{r colorblind friendly 3 Tritanomaly}

#Replication of code block above with one change: colors changed for accessibility to people with a blue-yellow colorblindness
ggplot(purifiedhoney, aes(year, prodsimpler)) +
  geom_line(aes(group=state),
            color = "gray80") +
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "#FF8F82") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "#00AE9C") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "#FFD0BC") +
  scale_y_log10()+
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "An adult worker bee's lifespan is about 45 days.")+    
  annotate(geom = "point", x = 1998, y = 37350, colour = "#FFD0BC", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "#00AE9C", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "#FF8F82", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")
```
\n\n 

Vizualization #5: Alteration for Accessibility #4 \n\n

```{r colorblind friendly desaturated}

#Replication of code block above with a few changes: colors completely desaturated for accessibility, and background gray altered (using theme_minimal)
ggplot(purifiedhoney, aes(year, prodsimpler)) +
  geom_line(aes(group=state),
            color = "gray95") +
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "gray10") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "gray65") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "gray40") +
  scale_y_log10()+
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "99% of a bee colony is made up of female bees.")+  
  theme_minimal()+
  annotate(geom = "point", x = 1998, y = 37350, colour = "gray40", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "gray70", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "gray10", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")
```

\n 

Vizualization #6: Alteration for Accessibility #5 (perhaps most accessible to people with any form of colorblindness) \n\n

```{r color blind safe palette}

#Replication of code block above, with colors selected specifically to be safe with an audience that is colorblind
ggplot(purifiedhoney, aes(year, prodsimpler)) +
  geom_line(aes(group=state),
            color = "gray90") +
  geom_line(data = filter(purifiedhoney,
                           state == "OR"),
           color = "#00A078") +
    geom_line(data = filter(purifiedhoney,
                           state == "WA"),
           color = "#FC6426") +
  
    geom_line(data = filter(purifiedhoney,
                           state == "CA"),
           color = "#FFA12C") +
  scale_y_log10()+
  labs(title = "US Honey Production",
       subtitle = "1998-2012",
       y = "Thousands of pounds",
       x = "Year",
       caption = "Bees have an excellent sense of smell.")+    
  theme_minimal()+
  annotate(geom = "point", x = 1998, y = 37350, colour = "#FFA12C", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 37350, label = "CALIFORNIA", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2955, colour = "#FC6426", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 3500, label = "WASHINGTON", hjust = "left")+    
  annotate(geom = "point", x = 1998, y = 2250, colour = "#00A078", size = 3) + 
  annotate(geom = "text", x = 1998.3, y = 1900, label = "OREGON", hjust = "left")

```

```{r join with census regions}

#Join for existing dataset (purifiedhoney) and additional data for assignment from TidyTuesday author (census_regions). Key is automatically identified as "state"
regional_honey_production <- left_join(purifiedhoney, census_regions) %>% 
  group_by(state, Region) %>% 
  summarise(totalprod = mean(totalprod)) %>% 
  ungroup() %>% 
  mutate(Region = as.factor(Region))
```

\n Vizualization #7: Bar Plot of Average Honey Production by Region \n\n

```{r bar plot showing average honey production, with color indicating region}

#it's all in the title to this code chunk
ggplot(regional_honey_production, aes(reorder(state, -totalprod), totalprod, fill = Region)) +
  geom_col(aes(group = Region)) +
  coord_flip() +
  #removed exponential labels along the x-axis
   scale_y_continuous(labels = comma) +
  labs(title = "Average Honey Production 1998-2012",
       subtitle = "Per state with color indicating region", 
       y = "Average annual production in gallons",
       x = "State")

```

Most honey production for the United States occurs in the Midwest and West. This is not surprising as most of the crop production for the country also occurs in these areas; bees would thrive in areas where there are high levels of crop production. The lowest production rate is in the northeast. \n\n

\n Vizualization #8: Heatmap of Honey Production by Region \n\n

```{r heatmap of average production by region}

premium_honey <- imported_honey_production %>%
  select(state, year)

# Data for heatmap with mean calculations embedded 
heat_dat <- full_join(premium_honey, regional_honey_production) %>%
  group_by(year, Region) %>% 
  mutate(totalprod = round(mean(totalprod)))

# Heatmap
ggplot(heat_dat, aes(year, Region, fill = totalprod)) +
  geom_raster(hjust = 0, vjust = 0) +
  scale_fill_viridis_c(option = "magma") +
  labs(title = "Average Honey Production by Region", 
       subtitles = "1998-2012",
       x = "Year",
       y = "Region")

```

\n Vizualization #9: Choice Plot! Made to resemble a honeycomb  \n\n

```{r choice plot}

#For our choice plot, we elected to do a honeycomb type figure
ggplot(imported_honey_production, aes(year, priceperlb)) +
  geom_hex(bins = 12) +
  scale_y_continuous(labels = dollar_format()) +
  #We liked a previous version of our code with different colors, but given the parameters of the assignment it might be more appropriate to have a differential based on shade rather than color, so we didn't force scale_fill_viridis below
  #scale_fill_viridis_c() +
  theme_minimal() +
  #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  labs(title = "Honey Prices per Pound from 1998 to 2012",
       subtitle = "Per state",
       x = "Year",
       y = "Price per pound",
       caption = "Plot labels reflect select year average price per pound.") +
  scale_fill_gradient(low = "#FFFC75", high = "#FF942A", name = "Number of States per unit") +
  
  # Tried to add in labels for yearly average
  
  geom_text(data = yr_means, aes(x=year, y=avg, label=avg))

# Quick data frame for yearly average labels
yr_means <- imported_honey_production %>% 
  group_by(year) %>% 
  summarize(round(mean(priceperlb), digits = 2)) %>% 
  rename(avg = "round(mean(priceperlb), digits = 2)") %>% 
  filter(year == 2000 |
         year == 2004 |
         year == 2008 |
         year == 2012)

# I hoped to include dollar signs in the plot labels through geom_text but wasn't able to
# figure out how to do so.

```